Multilevel generalizations of collective spin dynamics and squeezing

### 2019-10-25 collective_sun/multilevel_spin_notes.tex
added content to the section in the main text about perturbative SOC introduced a scaling factor for drive operators

We consider a collection of $n$-level fermions on a 1-D lattice, with
a single-particle Hamiltonian
\begin{align}
  H_{\t{lat}} = -t \sum_{q,\mu} \cos\p{qa} c_{q\mu}^\dag c_{q\mu},
  \label{eq:H_lat}
\end{align}
where $t$ is the nearest-neighbor tunneling rate, $a$ is the spacing
between neighboring lattice sites, and $c_{q\mu}$ is the fermionic
annihilation operator for a particle in quasi-momentum mode $q$ with
internal state (e.g.~nuclear spin) $\mu$.  Under the frozen-mode
approximation, collective two-body SU($n$)-symmetric interactions
between these fermions take form
\begin{align}
  H_{\t{int}} = \f{u}{2} \sum_{p,q,\mu,\nu}
  \p{c_{p\mu}^\dag c_{q\nu}^\dag c_{q\nu} c_{p\mu}
    + c_{q\mu}^\dag c_{p\nu}^\dag c_{q\nu} c_{p\mu}},
\end{align}
where $u\equiv U/L$ is the two-body on-site interaction energy $U$
divided by the number of lattice sites $L$.  Defining the spin
operators $S_{\mu\nu}^{(p)}\equiv c_{p\mu}^\dag c_{p\nu}$ for each
quasi-momentum $p$, we can alternately write
\begin{align}
  H_{\t{int}} = \f{u}{2} \sum_{p,q,\mu,\nu}
  \p{S_{\mu\mu}^{(p)} S_{\nu\nu}^{(q)} - S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)}}
  = \f{u}{2}\p{N^2 - \sum_{p,q,\mu,\nu}
    S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)}},
\end{align}
where $N$ is the total number of particles.  In the absence of
coherence between sectors of different particle number, we can neglect
the $\sim N^2$ term and simply write
\begin{align}
  H_{\t{int}}
  = - \f{u}{2} \sum_{p,q,\mu,\nu} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)}.
\end{align}
In the case of SU(2), at this point we would expand the spin operators
$S_{\mu\nu}^{(p)}$ in terms of Pauli operators in order to convert
$H_{\t{int}}$ into an SU(2) spin Hamiltonian\cite{he2019engineering}.
To generalize this procedure to the case of SU($n$), we first define a
vector $\v S^{(p)}$ of all spin operators $S_{\mu\nu}^{(p)}$, and
write
\begin{align}
  H_{\t{int}}
  = - \f{u}{2} \sum_{p,q} {\v S^{(p)}}^\dag \c \v S^{(q)}
  = - \f{u}{2} \v\S^\dag \c \v\S,
  &&
  \v\S\equiv \sum_p \v S^{(p)}.
  \label{eq:H_int_spin}
\end{align}
In the bulk of this section, we identify various bases for the space
of linear operators on the Hilbert space $\H_n$ of an $n$-level
system.  For brevity, we denote this space of operators by
$\B\p{\H_n}$, and note that $\B\p{\H_n}$ is itself an
$n^2$-dimensional Hilbert space equipped with a trace
(Hilbert-Schmidt) inner product
\begin{align}
  \obk{\O|\Q} \equiv \tr\p{\O^\dag Q},
\end{align}
where $\oket{\Q}$ and $\obra{\O}$ respectively denote vectors in
$\B\p{\H_n}$ and its dual space $\B\p{\H_n}^*$.  Similarly to the set
of all spin operators $S_{\mu\nu}\equiv\op{\mu}{\nu}$ on the
single-particle Hilbert space at each quasi-momentum, all of the bases
we consider will be orthonormal with respect to the trace inner
product.  As we show in Appendix \ref{sec:changing_bases}, the
collective SU($n$)-symmetric interaction Hamiltonian in
\eqref{eq:H_int_spin} takes an identical form in any orthonormal basis
for $\B\p{\H_n}$.

@@

The drive operators $D_{1,0}$ and $D_{1,\pm1}$ are proportional to the
Hamiltonians induced on an $n$-level spin by an external magnetic
field and a classical spin-polarized driving laser; loosely speaking,
one can identify $\p{D_{1,0},D_{1,1},D_{1,-1}}\sim\p{S_\z,S_\x,S_\y}$,
where $S_\alpha$ is a collective SU(2) spin-$\alpha$ operator for
$\p{n-1}$ 2-level spins\cite{perlin2019shorttime}.  More concretely,
we can embed the Hilbert space of a single $n$-level spin into the
permutationally-symmetric (Dicke) manifold of states for $\p{n-1}$
2-level spins via
\begin{align}
  \ket{\mu}_{\t{single}} \to \ket{I+\mu}_{\t{collective}}
  \propto S_+^{I+\mu} \ket\dn^{\otimes\p{n-1}},
\end{align}
where $I\equiv\p{n-1}/2$ is a total spin,
$\mu\in\set{-I,-I+1,\cdots,I}$ indexes a state of the single $n$-level
spin, and $S_+$ is a collective spin-raising operator for $\p{n-1}$
2-level spins.  This embedding takes
\begin{align}
  \p{D_{1,0},D_{1,1},D_{1,-1}}
  \to \p{S_\z,S_\x,S_\y}
  \times \f{2\sqrt{3}}{\sqrt{\p{n-1}n\p{n+1}}}.
  \label{eq:spin_ops}
\end{align}
Furthermore, if we {\it define} the operators $S_\alpha$ for a single
$n$-level spin via the correspondence in \eqref{eq:spin_ops}, then the
familiar SU(2) transformation rule
\begin{align}
  e^{i\phi S_\z} S_\x e^{-i\phi S_\z} = S_\x^{(\phi)},
  &&
  S_\x^{(\phi)} \equiv \cos\phi S_\x - \sin\phi S_\y,
\end{align}
generalizes to
\begin{align}
  e^{i\phi S_\z} D_{LM} e^{-i\phi S_\z} = D_{LM}^{(\phi)},
  &&
  D_{LM}^{(\phi)} \equiv \cos\p{M\phi} D_{LM} - \sin\p{M\phi} D_{L,-M}.
\end{align}

@@

For our final re-expression of collective SU($n$)-symmetric
interactions, we abandon the use of single-spin operators entirely to
express the interaction Hamiltonian $H_{\t{int}}$ in terms of
operators that permute tensor factors (subsystems) of a Hilbert space
with an $N$-fold tensor product structure (e.g.~an array of nuclear
spins).  Specifically, we denote the Hilbert space of an $n$-level
system by $\H_n$, denote its $N$-fold tensor product by
$\H_n^{\otimes N}$, and identify the set $\Pi\equiv\set{\Pi_\sigma}$
of operators that permute tensor factors of $\H_n^{\otimes N}$
according to elements $\sigma\in\SS_N$ of the symmetric group of order
$N$.  We then expand
\begin{align}
  H_{\t{int}}
  = -\f{u}{2} \sum_{p,q,\mu,\nu} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)}
  = -u\sum_{\substack{p<q\\\mu,\nu}} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)}
  - \f{u}{2} \sum_{p,\mu,\nu} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(p)},
  \label{eq:H_int_perm_start}
\end{align}
where the first sum,
\begin{align}
  \sum_{\substack{p<q\\\mu,\nu}} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)}
  = \sum_{\substack{p<q\\\mu,\nu}} \op{\mu}{\nu}^{(p)} \op{\nu}{\mu}^{(q)}
  = \sum_{\substack{p<q\\\mu,\nu}} \op{\mu\nu}{\nu\mu}^{(p,q)}
  = \sum_{p<q} \Pi_{pq}
  \label{eq:spin_perm}
\end{align}
is simply a sum over all permutations $\Pi_{pq}$ of subsystems $p$ and
$q$; and
\begin{align}
  \sum_{p,\mu,\nu} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(p)}
  = \sum_{p,\mu,\nu} \op{\mu}{\nu}^{(p)} \op{\nu}{\mu}^{(p)}
  = n \sum_{p,\mu} \op{\mu}{\mu}^{(p)}
  = n \sum_p \1^{(p)}
  = n N
  \label{eq:spin_const}
\end{align}
is merely a constant.  Substituting \eqref{eq:spin_perm} and
\eqref{eq:spin_const} into \eqref{eq:H_int_perm_start} yields
\begin{align}
  H_{\t{int}} = -u \sum_{p<q} \Pi_{pq} - \f12 N n u
  \simeq -u\sum_{p<q} \Pi_{pq},
  \label{eq:H_int_perm}
\end{align}
where $\simeq$ denotes equality up to an overall constant with no
physical consequence.  The form of the interaction Hamiltonian
$H_{\t{int}}$ makes it evident that its ground-state manifold
$\M_{\FS}$ is the set of {\it fully symmetric} (FS) states that have
eigenvalue $1$ with respect to all pairwise permutation operators
$\Pi_{pq}$, and thereby also all $\Pi_\sigma\in\Pi$.  The fully
symmetric manifold $\M_\FS$ is spanned by a basis $\set{\ket{\v m}}$
of fully symmetric states that can be uniquely identified by a list of
integers $\v m\equiv\p{m_0,m_1,\cdots,m_{n-1}}\in\ZZ_{N+1}^n$, where
$m_s$ indicates the total occupation number of subsystem state $s$,
and $\sum_sm_s=N$.  In the case of $n=2$, the fully symmetric manifold
$\M_\FS$ is precisely the Dicke manifold\cite{dicke1954coherence} of
$N$ qubits, spanned by Dicke states $\set{\ket{m_0,m_1}}$ for which
$m_0+m_1=N$ and the integer $m_0$ ($m_1$) indicates the total
occupation number of qubit state $\ket{0}$ ($\ket{1}$).  The dimension
of fully symmetric manifold $\M_\FS$ is determined by the number of
ways to assign $N$ subsystems to $n$ states, which is
${N+n-1 \choose n-1}$, and the energy of a state
$\ket{\psi_\FS}\in\M_\FS$ with respect to the interaction Hamiltonian
$H_{\t{int}}$ in \eqref{eq:H_int_perm} is
$-u{N \choose 2}=-\p{1/4}N\p{N-1}u$.

@@

A driving laser addressing spin-$1/2$ fermions on a 1-D lattice will
induce the Hamiltonian
\begin{align}
  \left. H_{\t{drive}}^{(\phi)} \right|_{n=2}
  = \Omega \sum_j
  \p{e^{i\phi j} c_{j,\up}^\dag c_{j,\dn}
    + e^{-i\phi j} c_{j,\dn}^\dag c_{j,\up}},
  \label{eq:H_drive_2}
\end{align}
where $\Omega$ is a (real) driving amplitude, $j\in\mathbb{Z}$ indexes
a lattice site, and $\phi$ is the relative phase of the drive between
adjacent lattice sites.  The multilevel analogue of
\eqref{eq:H_drive_2} is
\begin{align}
  H_{\t{drive}}^{(\phi)}
  = -\Omega \sum_j \p{e^{i\phi j} T_{1,1}^{(j)}
    + e^{-i\phi j} {T_{1,1}^{(j)}}^\dag}
  = \Omega \D_{1,1}^{(\phi)},
  \label{eq:H_drive}
\end{align}
where
\begin{align}
  \D_{LM}^{(\phi)}
  \equiv \sum_j \sp{\cos\p{M\phi j} D_{LM}^{(j)}
    - \sin\p{M\phi j} D_{L,-M}^{(j)}},
  \label{eq:drive_rot}
\end{align}
is an inhomogeneously rotated collective drive operator.  We can make
this drive spatially homogenous via the gauge transformation
$c_{j\mu}^\dag \to e^{-i\mu\phi j} c_{j\mu}^\dag$, which takes
 \begin{align}
  H_{\t{drive}}^{(\phi)} \to H_{\t{drive}} = \Omega \D_{1,1},
  &&
  \D_{LM} \equiv \D_{LM}^{(0)} = \sum_j D_{LM}^{(j)}.
\end{align}
The SU($n$) symmetry of collective interactions considered in this
work implies that the interaction Hamiltonian $H_{\t{int}}$ in
\eqref{eq:H_int_perm} is unaffected by this gauge transformation.  The
single-particle Hamiltonian $H_{\t{lat}}$ in \eqref{eq:H_lat},
however, transforms as (see Appendix \ref{sec:lat_drive})
\begin{align}
  H_{\t{lat}}
  \to H_{\t{lat}}^{(\phi)}
  = -t \sum_{q,\mu} \cos\p{qa+\mu\phi} c_{q\mu}^\dag c_{q\mu}
  = \sum_{q,L} B_L^{(\phi,q)} D_{L,0}^{(q)},
  \label{eq:H_lat_SOC_B}
\end{align}
with effective inhomogeneous driving field amplitudes
\begin{align}
  B_L^{(\phi,q)} = -t w_L\p{qa} A_L^{(\phi)},
\end{align}
where
\begin{align}
  w_L\p{\theta} \equiv
  \begin{cases}
    \cos\theta & L~\t{even} \\
    \sin\theta & L~\t{odd}
  \end{cases},
  &&
  A_L^{(\phi)} \equiv \p{-1}^L \sqrt{\f{2L+1}{2I+1}}
  \sum_\mu \bk{I\mu;L,0|I\mu} w_L\p{\mu\phi}.
  \label{eq:A_L_phi}
\end{align}
The multilevel spin-orbit coupling (SOC) exhibited by the
singe-particle Hamiltonian $H_{\t{lat}}^{(\phi)}$ thus allows for the
simulation of synthetic gauge fields with large spin ($L>1$).  As we
will see in Section \ref{sec:drive_raman}, multi-laser drives can also
simulate synthetic large-spin fields with non-zero spin projection
($M\ne0$) onto a quantization axis.

@@

The single-particle Hamiltonian $H_{\t{lat}}^{(\phi)}$ breaks
permutational symmetry, thereby coupling states within the fully
symmetric manifold $\M_\FS$ to states outsides it.  If SOC is
sufficiently weak, however, transitions outside the fully symmetric
manifold $\M_\FS$ are energetically suppressed by interactions, and we
can account for the effect of SOC perturbatively.  In the regime of
weak SOC with $\abs{\phi}\ll1$, by considering
\begin{enumerate*}
\item all $L\le12$ for arbitrary $n$, and
\item all $L<n$ for all $n\le25$
\end{enumerate*}
(and conjecturing the same functional form for all $L$ and $n$), we
find that
% expand Clebsch-Gordan coefficients in a power series according to
% \url{functions.wolfram.com/07.38.06.0008.01}
\begin{align}
  A_L^{(\phi)} = c_L \sqrt{n_L}\, \phi^L + O\p{\phi^{L+2}},
  &&
  c_L \equiv \p{-1}^{L+\lfloor L/2\rfloor}
  \f{\sqrt{2L+1}\,L!}{\p{2L+1}!},
  &&
  n_L \equiv \prod_{\ell=-L}^L\p{n+\ell}.
  \label{eq:A_L_phi_small}
\end{align}

\vspace{3cm}

@@

adjacent lattice sites; $\op{r}{s}$ for $r,s\in\set{\up,\dn}$ is a

@@

  {T_{1,m}^{(j)}}^\dag T_{1,m}^{(j)}, & \label{eq:raman_0} \\

@@

  \Omega_{LM}^{(\phi,j)} D_{LM}^{(j)},

@@

  = \sum_{\substack{L\le2\\0\le M\le L}} \Omega_{LM} \D_{LM}^{(\phi)},

@@

  = \sum_{\substack{L\le2\\0\le M\le L}} \Omega_{LM} \D_{LM}.

@@

where $f\equiv N/L$ is the filling fraction of quasi-momentum modes.
We have thus far simplified the effective Hamiltonian in
\eqref{eq:H_eff_2_start} to

>>>>

We consider a collection of $n$-level fermions on a 1-D lattice, with
a single-particle Hamiltonian
\begin{align}
  H_{\t{lat}} = -t \sum_{q,\mu} \cos\p{qa} c_{q\mu}^\dag c_{q\mu},
  \label{eq:H_lat}
\end{align}
where $t$ is the nearest-neighbor tunneling rate, $a$ is the spacing
between neighboring lattice sites, and $c_{q\mu}$ is the fermionic
annihilation operator for a particle in quasi-momentum mode $q$ with
internal state (e.g.~nuclear spin) $\mu$.  Under the frozen-mode
approximation, collective two-body SU($n$)-symmetric interactions
between these fermions take form
\begin{align}
  H_{\t{int}} = \f{u}{2} \sum_{p,q,\mu,\nu}
  \p{c_{p\mu}^\dag c_{q\nu}^\dag c_{q\nu} c_{p\mu}
    + c_{q\mu}^\dag c_{p\nu}^\dag c_{q\nu} c_{p\mu}},
\end{align}
where $u\equiv fU/N$ with $U$ the two-body on-site interaction energy,
$f$ the filling fraction of lattice sites, and $N$ the total number of
fermions.  Defining the spin operators
$S_{\mu\nu}^{(p)}\equiv c_{p\mu}^\dag c_{p\nu}$ for each
quasi-momentum $p$, we can alternately write
\begin{align}
  H_{\t{int}} = \f{u}{2} \sum_{p,q,\mu,\nu}
  \p{S_{\mu\mu}^{(p)} S_{\nu\nu}^{(q)} - S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)}}
  = \f{u}{2}\p{N^2 - \sum_{p,q,\mu,\nu}
    S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)}},
\end{align}
where $N$ is the total number of particles.  In the absence of
coherence between sectors of different particle number, we can neglect
the $\sim N^2$ term and simply write
\begin{align}
  H_{\t{int}}
  = - \f{u}{2} \sum_{p,q,\mu,\nu} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)}.
\end{align}
In the case of SU(2), at this point we would expand the spin operators
$S_{\mu\nu}^{(p)}$ in terms of Pauli operators in order to convert
$H_{\t{int}}$ into an SU(2) spin Hamiltonian\cite{he2019engineering}.
To generalize this procedure to the case of SU($n$), we first define a
vector $\v S^{(p)}$ of all spin operators $S_{\mu\nu}^{(p)}$, and
write
\begin{align}
  H_{\t{int}}
  = - \f{u}{2} \sum_{p,q} {\v S^{(p)}}^\dag \c \v S^{(q)}
  = - \f{u}{2} \v\S^\dag \c \v\S,
  &&
  \v\S\equiv \sum_p \v S^{(p)}.
  \label{eq:H_int_spin}
\end{align}
In the bulk of this section, we identify various bases for the space
of linear operators on the Hilbert space $\H_n$ of an $n$-level
system.  For brevity, we denote this space of operators by
$\B\p{\H_n}$, and note that $\B\p{\H_n}$ is itself an
$n^2$-dimensional Hilbert space equipped with a trace
(Hilbert-Schmidt) inner product
\begin{align}
  \obk{\O|\Q} \equiv \tr\p{\O^\dag Q},
\end{align}
where $\oket{\Q}$ and $\obra{\O}$ respectively denote vectors in
$\B\p{\H_n}$ and its dual space $\B\p{\H_n}^*$.  Similarly to the set
of all spin operators $S_{\mu\nu}\equiv\op{\mu}{\nu}$ on the
single-particle Hilbert space at each quasi-momentum, all of the bases
we consider will be orthonormal with respect to the trace inner
product.  As we show in Appendix \ref{sec:changing_bases}, the
collective SU($n$)-symmetric interaction Hamiltonian in
\eqref{eq:H_int_spin} takes an identical form in any orthonormal basis
for $\B\p{\H_n}$.

@@

The drive operators $D_{1,0}$ and $D_{1,\pm1}$ are proportional to the
Hamiltonians induced on an $n$-level spin by an external magnetic
field and a classical spin-polarized driving laser; loosely speaking,
one can associate
$\p{D_{1,0},D_{1,1},D_{1,-1}}\sim\p{S_\z,S_\x,S_\y}$, where $S_\alpha$
is a collective SU(2) spin-$\alpha$ operator for $\p{n-1}$ 2-level
spins\cite{perlin2019shorttime}.  More concretely, we can embed the
Hilbert space of a single $n$-level spin into the
permutationally-symmetric (Dicke) manifold of states for $\p{n-1}$
2-level spins via
\begin{align}
  \ket{\mu}_{\t{single}} \to \ket{I+\mu}_{\t{collective}}
  \propto S_+^{I+\mu} \ket\dn^{\otimes\p{n-1}},
  \label{eq:collective_embedding}
\end{align}
where $I\equiv\p{n-1}/2$ is a total spin,
$\mu\in\set{-I,-I+1,\cdots,I}$ indexes a state of the single $n$-level
spin, and $S_+$ is a collective spin-raising operator for $\p{n-1}$
2-level spins.  This embedding takes
\begin{align}
  \xi_1 \times \p{D_{1,0}, D_{1,1}, D_{1,-1}}
  \to \p{S_\z, S_\x, S_\y},
  \label{eq:spin_ops}
\end{align}
where
\begin{align}
  \xi_L
  \equiv \sqrt{2L+1}\, \f{L!}{\p{2L+1}!}
  \sp{\prod_{\ell=-L}^L\p{n+\ell}}^{1/2},
  \label{eq:scale_fac}
\end{align}
is a scale factor that will appear frequently in the theory of drive
operators.

@@

For our final re-expression of collective SU($n$)-symmetric
interactions, we abandon the use of single-spin operators entirely to
express the interaction Hamiltonian $H_{\t{int}}$ in terms of
operators that permute tensor factors (subsystems) of a Hilbert space
with an $N$-fold tensor product structure (e.g.~an array of nuclear
spins).  Specifically, we denote the Hilbert space of an $n$-level
system by $\H_n$, denote its $N$-fold tensor product by
$\H_n^{\otimes N}$, and identify the set $\Pi\equiv\set{\Pi_\sigma}$
of operators that permute tensor factors of $\H_n^{\otimes N}$
according to elements $\sigma\in\SS_N$ of the symmetric group of order
$N$.  We then expand
\begin{align}
  H_{\t{int}}
  = -\f{u}{2} \sum_{p,q,\mu,\nu} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)}
  = -u\sum_{\substack{p<q\\\mu,\nu}} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)}
  - \f{u}{2} \sum_{p,\mu,\nu} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(p)},
  \label{eq:H_int_perm_start}
\end{align}
where the first sum,
\begin{align}
  \sum_{\substack{p<q\\\mu,\nu}} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(q)}
  = \sum_{\substack{p<q\\\mu,\nu}} \op{\mu}{\nu}^{(p)} \op{\nu}{\mu}^{(q)}
  = \sum_{\substack{p<q\\\mu,\nu}} \op{\mu\nu}{\nu\mu}^{(p,q)}
  = \sum_{p<q} \Pi_{pq}
  \label{eq:spin_perm}
\end{align}
is simply a sum over all permutations $\Pi_{pq}$ of subsystems $p$ and
$q$; and
\begin{align}
  \sum_{p,\mu,\nu} S_{\mu\nu}^{(p)} S_{\nu\mu}^{(p)}
  = \sum_{p,\mu,\nu} \op{\mu}{\nu}^{(p)} \op{\nu}{\mu}^{(p)}
  = n \sum_{p,\mu} \op{\mu}{\mu}^{(p)}
  = n \sum_p \1^{(p)}
  = n N
  \label{eq:spin_const}
\end{align}
is merely a constant.  Substituting \eqref{eq:spin_perm} and
\eqref{eq:spin_const} into \eqref{eq:H_int_perm_start} yields
\begin{align}
  H_{\t{int}} = -u \sum_{p<q} \Pi_{pq} - \f12 N n u
  \simeq -u\sum_{p<q} \Pi_{pq},
  \label{eq:H_int_perm}
\end{align}
where $\simeq$ denotes equality up to an overall constant with no
physical consequence.  The form of the interaction Hamiltonian
$H_{\t{int}}$ makes it evident that its ground-state manifold
$\M_{\FS}$ is the set of {\it fully symmetric} (FS) states that have
eigenvalue 1 with respect to all pairwise permutation operators
$\Pi_{pq}$, and thereby also all $\Pi_\sigma\in\Pi$.  The fully
symmetric manifold $\M_\FS$ is spanned by a basis $\set{\ket{\v m}}$
of fully symmetric states that can be uniquely identified by a list of
integers $\v m\equiv\p{m_0,m_1,\cdots,m_{n-1}}\in\ZZ_{N+1}^n$, where
$m_s$ indicates the total occupation number of subsystem state $s$,
and $\sum_sm_s=N$.  In the case of $n=2$, the fully symmetric manifold
$\M_\FS$ is precisely the Dicke manifold\cite{dicke1954coherence} of
$N$ qubits, spanned by Dicke states $\set{\ket{m_0,m_1}}$ for which
$m_0+m_1=N$ and the integer $m_0$ ($m_1$) indicates the total
occupation number of qubit state $\ket{0}$ ($\ket{1}$).  The dimension
of fully symmetric manifold $\M_\FS$ is determined by the number of
ways to assign $N$ subsystems to $n$ states, which is
${N+n-1 \choose n-1}$, and the energy of a state
$\ket{\psi_\FS}\in\M_\FS$ with respect to the interaction Hamiltonian
$H_{\t{int}}$ in \eqref{eq:H_int_perm} is
$-u{N \choose 2}=-\p{1/4}N\p{N-1}u$.

@@

A driving laser addressing spin-1/2 fermions on a 1-D lattice will
induce the Hamiltonian
\begin{align}
  \left. H_{\t{drive}}^{(\phi)} \right|_{n=2}
  = \f12 \Omega \sum_j
  \p{e^{i\phi j} c_{j,\up}^\dag c_{j,\dn}
    + e^{-i\phi j} c_{j,\dn}^\dag c_{j,\up}},
  \label{eq:H_drive_2}
\end{align}
where $\Omega$ is a (real) driving amplitude, $j\in\mathbb{Z}$ indexes
a lattice site, and $\phi$ is the relative phase of the drive between
adjacent lattice sites.  The multilevel analogue of
\eqref{eq:H_drive_2} is
\begin{align}
  H_{\t{drive}}^{(\phi)}
  = -\f12 \Omega \sum_j \p{e^{i\phi j} T_{1,1}^{(j)}
    + e^{-i\phi j} {T_{1,1}^{(j)}}^\dag}
  = \tilde \D_{1,1}^{(\phi)},
  &&
  \tilde\Omega \equiv \Omega \xi_1
  \label{eq:H_drive}
\end{align}
where
\begin{align}
  \D_{LM}^{(\phi)}
  \equiv \sum_j \sp{\cos\p{M\phi j} D_{LM}^{(j)}
    - \sin\p{M\phi j} D_{L,-M}^{(j)}},
  \label{eq:drive_rot}
\end{align}
is an inhomogeneously rotated collective drive operator.  We can make
this drive spatially homogenous via the gauge transformation
$c_{j\mu}^\dag \to e^{-i\mu\phi j} c_{j\mu}^\dag$, which takes
 \begin{align}
   H_{\t{drive}}^{(\phi)} \to H_{\t{drive}} = \tilde\Omega \D_{1,1},
   &&
   \D_{LM} \equiv \D_{LM}^{(0)} = \sum_j D_{LM}^{(j)}.
\end{align}
The SU($n$) symmetry of collective interactions considered in this
work implies that the interaction Hamiltonian $H_{\t{int}}$ in
\eqref{eq:H_int_perm} is unaffected by this gauge transformation.  The
single-particle Hamiltonian $H_{\t{lat}}$ in \eqref{eq:H_lat},
however, transforms as (see Appendix \ref{sec:lat_drive})
\begin{align}
  H_{\t{lat}}
  \to H_{\t{lat}}^{(\phi)}
  = -t \sum_{q,\mu} \cos\p{qa+\mu\phi} c_{q\mu}^\dag c_{q\mu}
  = \sum_{q,L} B_L^{(\phi,q)} D_{L,0}^{(q)},
  \label{eq:H_lat_SOC_B}
\end{align}
with effective inhomogeneous driving field amplitudes
\begin{align}
  B_L^{(\phi,q)} = -t w_L\p{qa} A_L^{(\phi)},
\end{align}
where
\begin{align}
  w_L\p{\theta} \equiv
  \begin{cases}
    \cos\theta & L~\t{even} \\
    \sin\theta & L~\t{odd}
  \end{cases},
  &&
  A_L^{(\phi)} \equiv \p{-1}^L \sqrt{\f{2L+1}{2I+1}}
  \sum_\mu \bk{I\mu;L,0|I\mu} w_L\p{\mu\phi}.
  \label{eq:A_L_phi}
\end{align}
The multilevel spin-orbit coupling (SOC) exhibited by the
singe-particle Hamiltonian $H_{\t{lat}}^{(\phi)}$ thus allows for the
simulation of synthetic gauge fields with large spin ($L>1$).  As we
will see in Section \ref{sec:drive_raman}, multi-laser drives can also
simulate synthetic large-spin fields with non-zero spin projection
($M\ne0$) onto a quantization axis.

@@

The single-particle Hamiltonian $H_{\t{lat}}^{(\phi)}$ breaks
permutational symmetry, thereby coupling states within the fully
symmetric manifold $\M_\FS$ to states outsides it.  If SOC is
sufficiently weak, however, transitions outside the fully symmetric
manifold $\M_\FS$ are energetically suppressed by interactions, and we
can account for the effect of SOC perturbatively.  In the regime of
weak SOC with $\abs{\phi}\ll1$, by considering
\begin{enumerate*}
\item all $L\le12$ for arbitrary $n$, and
\item all $L<n$ for all $n\le25$
\end{enumerate*}
(and conjecturing the same functional form for all $L$ and $n$), we
find that
% expand Clebsch-Gordan coefficients in a power series according to
% \url{functions.wolfram.com/07.38.06.0008.01}
\begin{align}
  A_L^{(\phi)}
  = \p{-1}^{L+\lfloor L/2\rfloor} \xi_L \phi^L + O\p{\phi^{L+2}},
  \label{eq:A_L_phi_small}
\end{align}
where $\xi_L$ is a scale factor defined in \eqref{eq:scale_fac}.  A
perturbative treatment of SOC through second order in the SOC angle
$\phi$ yields the following effective Hamiltonian in the fully
symmetric manifold $\M_\FS$ (see Appendix \ref{sec:SOC_pert}):
\begin{align}
  H_{\t{eff}} = H_{\t{eff}}^{(1)} + H_{\t{eff}}^{(2)},
\end{align}
\begin{align}
  H_{\t{eff}}^{(1)}
  &= \phi t \, \EE_p\sp{\sin\p{pa}} \xi_1 \D_{1,0}
  + \phi^2 t \, \EE_q\sp{\cos\p{qa}} \xi_2 \D_{2,0}
  + O\p{\phi^3},
  \label{eq:H_eff_1_phi} \\
  H_{\t{eff}}^{(2)}
  &= \f{\phi^2 t^2}{\p{N-1}fU} \var_p\sp{\sin\p{pa}}
  \p{\xi_1^2 \D_{1,0}^2 - 2 N \xi_2 \D_{2,0}}
  + O\p{\phi^3},
  \label{eq:H_eff_2_phi}
\end{align}
where
\begin{align}
  \EE_p\sp{X_p} \equiv \f1N \sum_p X_p,
  &&
  \var_p\sp{X_p} \equiv \EE_p\sp{\p{X_p-\EE_q\sp{X_q}}}
  = \EE_p\sp{X_p^2}-\EE_q\sp{X_q}^2,
\end{align}
are the mean and variance of $X_p$ over all occupied quasi-momenta
$p$.  In the case of SU(2), the effective Hamiltonians in
\eqref{eq:H_eff_1_phi} and \eqref{eq:H_eff_2_phi} reduce to the
spin-squeezing Hamiltonians derived in Ref.~\cite{he2019engineering}.
To facilitate comparison, we note that in the case of SU(2) the drive
operator $\D_{1,0}$ is precisely the collective spin-$z$ operator
$S_\z$ that appears in Ref.~\cite{he2019engineering}.  Furthermore,
drive operators $D_{LM}$ vanish when $L\ge n$, so in the case of SU(2)
the operator $\D_{2,0}$ has no contribution to the effective
Hamiltonians in \eqref{eq:H_eff_2_phi} and \eqref{eq:H_eff_2_phi}.

@@

adjacent lattice sites; we have added an overall factor of $\xi_1$ to
simplify later expressions; $\op{r}{s}$ for $r,s\in\set{\up,\dn}$ is a

@@

  {T_{1,m}^{(j)}}^\dag T_{1,m}^{(j)},
  \label{eq:raman_0} \\

@@

  \Omega_{LM}^{(\phi,j)} \xi_L D_{LM}^{(j)},

@@

  = \sum_{\substack{L\le2\\0\le M\le L}}
  \Omega_{LM} \xi_L \D_{LM}^{(\phi)},

@@

  = \sum_{\substack{L\le2\\0\le M\le L}} \Omega_{LM} \xi_L \D_{LM}.

@@

where $f$ is the filling fraction of lattice sites.  We have thus far
simplified the effective Hamiltonian in \eqref{eq:H_eff_2_start} to
